---
title: "OSM Can-BICS data quality model"
output:
  html_notebook: default
  word_document: default
---

## Background

The purpose of this project is to understand spatial variability in data quality 
for bike infrastructure in open data (from cities) and OpenStreetMap. 

* Infrastructure data are used to allocate infrastructure investments that can
improve safety for active transportation and increase active transportation 
rates.
* Data quality can influence the equity of infrastructure investments (good data
are needed to make good decisions), possibly disproportionately impacting safety 
for individuals in equity-seeking groups.
* Is data quality equitably distributed?
    * can we predict data quality (lack of data quality) using sociospatial 
    variables related to income and racial composition?
    * Are there spatial patterns in data quality? Do these relate to the spatial
    patterns of sociospatial variables?
* Can we create a model to identify priority locations for OSM contributions and
to inform fitness for use decisions for infrastructure data?

**Hypotheses**

1. Data quality for bike infrastructure is unevenly distributed in space.
2. Data quality is lower in places with spatial equity concerns.

## Methods

**Approach**

1. Explore relationships between sociospatial variables and bike infrastructure
reference data points.
    * Explore univariate relationships (graphs and maps).

  
2. Model data quality for reference points (n = 2000).
3. Model data quality all neighbourhoods in Canada using OSM intrinsic 
indicators of data quality as an outcome (# of mappers, # of edits).

**Analysis unit**

* In the exploratory analysis and the first model, we will apply a 100 m buffer 
to the sample points, since census boundaries are often on roads.
   * census data obtained for dissemination areas will be attributed to the 
   points using area weighted means.
* In the second model, we will apply a X km buffer to the population weighted 
centroids for dissemination areas (DAs), to provide within-city detail.
   * Neighbourhood sized.
   * Provides consistency with OSM Can-BICS.
   * DAs are the smallest standard unit and cover the entire country. They contain
a population between 400 to 700 people.
    * smaller where population density is higher.
    * very large rural DAs can make analyses difficult.
* Start in Vancouver and expand.


**Outcomes**

* The desired outcome is our ability to accurately predict Can-BICS class, 
relative to ground reference data both for:
   * open data, and
   * OSM Can-BICS.
   * Mapathon edits are another indicator of data quality (just Vancouver and 
   Edmonton have sufficient density to form hotspots). The presence of edits 
   indicate a lack of quality. Mapathon edits are lines - use the top 10% of 
   kernal density.
```{r, echo = F, include = F}
library(sf)
library(tidyverse)
library(mapview)
library(ggmap)
library(ggspatial)
library(spatstat) #  KDE
library(maptools)
library(raster)

# projections
geographic <- 4326
lambert <- 3347


source("../osm_infrastructure/code/Can_BICS_OSM_functions.R")

reference_points <- st_read("../osm_infrastructure/data/test_cities/reference_predicted.gpkg")
names(reference_points)

# reference
reference_points$Can_BICS_ground_class <- Can_BICS_class(reference_points$Can_BICS_ground)
reference_points$Can_BICS_ground_class <- reference_points$Can_BICS_ground_class %>%
  recode("None" = "Non-Conforming")
table(reference_points$Can_BICS_ground_class)

# open data
reference_points$Can_BICS_stratum_class <- Can_BICS_class(reference_points$Can_BICS_stratum)
reference_points$Can_BICS_stratum_class <- reference_points$Can_BICS_stratum_class %>%
  recode("None" = "Non-Conforming")
table(reference_points$Can_BICS_stratum_class)

# osm
reference_points$Can_BICS_class <- Can_BICS_class(reference_points$Can_BICS)
reference_points$Can_BICS_class <- reference_points$Can_BICS_class %>%
  recode("None" = "Non-Conforming")
table(reference_points$Can_BICS_class)

reference_points$match_osm <- reference_points$Can_BICS_ground_class == 
  reference_points$Can_BICS_class

reference_points$match_open_data <- reference_points$Can_BICS_ground_class == 
  reference_points$Can_BICS_stratum_class

reference_points_vancouver <- reference_points %>%
  filter(CSDNAME %in% "Vancouver")


st_write(reference_points_vancouver,
         "data/reference_predicted_vancouver.gpkg",
         delete_dsn = T)

```

```{r, echo=FALSE, message=FALSE}
  colour_scheme <- c("7570b3",
                     "#1b9e77")
  
  My_Theme = theme(
  axis.text.x=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks=element_blank(),
  panel.grid.major=element_blank(),
  panel.grid.minor=element_blank(),
  legend.position = "right")
  
  limits <- reference_points_vancouver %>% 
  st_bbox() %>%
  st_as_sfc() %>%
  st_buffer(100) %>%
  st_transform(4326) %>%
  st_bbox() %>%
  as.list()
  
  # geographic coords for map
  reference_points_vancouver_geo <- reference_points_vancouver %>%
    st_transform(4326)
  
  basemap <- get_stamenmap(bbox = c(left = limits$xmin,
                                    bottom = limits$ymin,
                                    right = limits$xmax,
                                    top = limits$ymax), 
                         maptype = "toner-background",
                         crop = T,
                         zoom = 12)
  
  
  match_open_data_map <- ggmap(basemap) +
  geom_sf(data = reference_points_vancouver_geo, 
          aes(col = match_open_data), 
          inherit.aes = FALSE,
          size = 2) + 
  scale_colour_manual(values = colour_scheme) +
  labs(col = "match_open_data") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)
  
match_open_data_map

```

```{r, echo = FALSE, message=FALSE}
  match_osm_data_map <- ggmap(basemap) +
  geom_sf(data = reference_points_vancouver_geo, 
          aes(col = match_osm), 
          inherit.aes = FALSE,
          size = 2) + 
  scale_colour_manual(values = c("purple", "green")) +
  labs(col = "match_osm") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

  match_osm_data_map
```
```{r, echo = FALSE, message=FALSE}
osm_edits <- st_read("data/mapathon/OSM_Edits.shp") %>%
  st_transform(3347)

  colour_scheme_comfort <- c("#238443",
                            "#c2e699",
                            "#2b8cbe",
                            "#969696")
  
  osm_edits_geo <- osm_edits %>%
    st_transform(4327)

  mapathon_map <- ggmap(basemap) +
    # inset_raster(as.raster(mapathon_edits.KDE.rast.geo),
    #              xmin = mapathon_edits.KDE.rast.geo@extent[1], 
    #              xmax = mapathon_edits.KDE.rast.geo@extent[2],
    #              ymin = mapathon_edits.KDE.rast.geo@extent[3], 
    #              ymax = mapathon_edits.KDE.rast.geo@extent[4]) +
    geom_sf(data = osm_edits_geo, 
          aes(col = comfort_cl), 
          inherit.aes = FALSE,
          size = 3) + 
  scale_colour_manual(values = colour_scheme_comfort) +
  labs(col = "Mapathon edits (comfort class)") +
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

  mapathon_map

```

```{r stats_for_mapathon}
library(knitr)
csds <- st_read("C:/Users/16043/Documents/basemap/census_2016/census_subdivision_boundary/lcsd000a16a_e.shp")

osm_edits_csds <- osm_edits %>% st_intersection(csds)
osm_edits_csds$length <- st_length(osm_edits_csds) %>% as.numeric()
osm_edits_csds$length <- round(osm_edits_csds$length / 1000, 0)

osm_edits_csds_summary <- osm_edits_csds %>%
  group_by(CSDNAME, PRNAME) %>%
  summarise(total_length = sum(length)) %>%
  filter(total_length > 0) %>% 
  arrange(PRNAME, CSDNAME, -total_length) %>%
  st_drop_geometry() %>%
  rename('CSD Name' = CSDNAME,
         'Province' = PRNAME,
         'length (km)' = total_length)

kable(osm_edits_csds_summary)

osm_edits_type_summary <- osm_edits_csds %>%
  group_by(infrastruc) %>%
  summarise(total_length = sum(length)) %>%
  filter(total_length > 0) %>% 
  arrange(-total_length) %>%
  st_drop_geometry() %>%
  rename('Infrastructure Type' = infrastruc,
         'length (km)' = total_length)

kable(osm_edits_type_summary)

```



```{r, include = FALSE}
# kde estimates
bandwidth <- 500
prank<-function(x) ifelse(is.na(x),NA,rank(x)/sum(!is.na(x)))
targetProjection <- crs("+proj=longlat +datum=WGS84 +no_defs")


window <- st_read("C:/Users/16043/Documents/basemap/census_2016/census_subdivision_cartographic/lcsd000b16a_e.shp") %>%
  filter(CSDNAME %in% "Vancouver") %>%
  st_transform(lambert) %>%
  dplyr::select(CSDNAME)

window.owin <- as.owin(as_Spatial(window))

# errors for open data
open_data_errors <- reference_points_vancouver %>%
  filter(! match_open_data) %>%
  dplyr::select(geom) %>%
  as.ppp()

Window(open_data_errors) <- window.owin

open_data_errors.KDE <- density.ppp(x = open_data_errors,
                                   kernel = "gaussian",
                                   sigma = bandwidth, 
                                   edge = F,
                                   eps = 10)

open_data_errors.KDE.rast <- raster(open_data_errors.KDE, crs = targetProjection)
open_data_errors.KDE.rast@data@values <- prank(open_data_errors.KDE.rast@data@values)

# errors for osm
osm_errors <- reference_points_vancouver %>%
  filter(! match_osm) %>%
  dplyr::select(geom) %>%
  as.ppp()

Window(osm_errors) <- window.owin

osm_errors.KDE <- density.ppp(x = osm_errors,
                                   kernel = "gaussian",
                                   sigma = bandwidth, 
                                   edge = F,
                                   eps = 10)

osm_errors.KDE.rast <- raster(osm_errors.KDE, crs = targetProjection)
osm_errors.KDE.rast@data@values <- prank(osm_errors.KDE.rast@data@values)

# mapathon edits
mapathon_edits <- osm_edits %>%
  mutate(count = 1) %>%
  dplyr::select(count) %>%
  as_Spatial() %>%
  as.psp()

Window(mapathon_edits) <- window.owin

mapathon_edits.KDE <- density.psp(x = mapathon_edits,
                                   kernel = "gaussian",
                                   sigma = bandwidth, 
                                   edge = F,
                                   eps = 10)

mapathon_edits.KDE.rast <- raster(mapathon_edits.KDE, crs = targetProjection)
mapathon_edits.KDE.rast@data@values <- prank(mapathon_edits.KDE.rast@data@values)
plot(mapathon_edits.KDE.rast)

```


* NEXT STEPS: Ground reference data is not available outside of the 15 sample cities. Where 
there is no ground reference data, we will use intrinsic indicators of data 
quality from OSM, such as the number of contributions, the number of 
contributors as a proxy.
   * in the first model, this will be used as a predictor to evaluate the 
   relationship with the desired outcome variable, and this will guide our
   selection of a suitable proxy variable.
   * in the second model, this will be an outcome variable.


**Predictor Variables**

***Canada Census Variables***

```{r, echo=F, message=FALSE}
#geometry
das <- st_read("C://Users/16043/Documents/basemap/census_2016/census_da_carto/lda_000b16a_e.shp") 


library(cancensus)
options(cancensus.api_key='CensusMapper_e0c23a6719376c837a4ae87e9e9227f2')
options(cancensus.cache_path = 'C:/working/cancensus')

vectors <- list_census_vectors("CA16")
#print(vectors[vectors$vector %in% "v_CA16_5690",])

census_data <- get_census(dataset='CA16',
                          regions=list(PR=c("59","48","47","46","35","24","10",
                                            "13","12","60","61","62")), 
                          vectors=c(pop_dens = "v_CA16_406",
                                    med_hous_inc = "v_CA16_2398",
                                    vis_min_tot = "v_CA16_3954",
                                    vis_min_pop = "v_CA16_3957",
                                    non_vis_min_pop = "v_CA16_3996",
                                    edu_tot = "v_CA16_5096",
                                    edu_seconary = "v_CA16_5102",
                                    edu_post_sec = "v_CA16_5105",
                                    edu_major_tot = "v_CA16_5141",
                                    edu_major_infotech = "v_CA16_5243",
                                    ocu_tot = "v_CA16_5654",
                                    ocu_na = "v_CA16_5657",
                                    ocu_all = "v_CA16_5660",
                                    ocu_mgmt = "v_CA16_5663",
                                    ocu_busi = "v_CA16_5666",
                                    ocu_sci = "v_CA16_5669",
                                    ocu_health = "v_CA16_5672",
                                    ocu_edu = "v_CA16_5675",
                                    ocu_art = "v_CA16_5678",
                                    ocu_sales = "v_CA16_5681",
                                    ocu_trades = "v_CA16_5684",
                                    ocu_natres = "v_CA16_5687",
                                    ocu_man = "v_CA16_5690",
                                    abor_tot = "v_CA16_3852",
                                    abor_ident = "v_CA16_3855",
                                    jtw_tot = "v_CA16_5792",
                                    jtw_bike = "v_CA16_5807"),
                          labels="short", 
                          geo_format=NA, 
                          level='DA')

census_data_clean <- census_data %>%
  mutate(bike_mode_share = round((100 * jtw_bike/jtw_tot), 0),
         vis_min_perc = round((100 * vis_min_pop/vis_min_tot), 0),
         non_vis_min_perc = round((100 * non_vis_min_pop/vis_min_tot), 0),
         ind_perc = round((100 * abor_ident/abor_tot), 0),
         edu_post_sec_perc = round((100 * edu_post_sec/edu_tot), 0),
         edu_post_info_tech_perc = round((100 * edu_major_infotech/edu_tot), 0),
         ocu_sci_perc = round((100 * ocu_sci/ocu_tot), 0),
         ocu_mgmt_perc = round((100 * ocu_mgmt/ocu_tot), 0), 
         ocu_busi_perc = round((100 * ocu_busi/ocu_tot), 0), 
         ocu_health_perc = round((100 * ocu_health/ocu_tot), 0), 
         ocu_edu_perc = round((100 * ocu_edu/ocu_tot), 0), 
         ocu_art_perc = round((100 * ocu_art/ocu_tot), 0), 
         ocu_trades_perc = round((100 * ocu_trades/ocu_tot), 0),
         ocu_natres_perc = round((100 * ocu_natres/ocu_tot), 0), 
         ocu_man_perc = round((100 * ocu_man/ocu_tot), 0)) %>%
  dplyr::select(GeoUID,
                med_hous_inc,
                pop_dens,
                med_hous_inc,
                bike_mode_share,
                vis_min_perc,
                ind_perc,
                edu_post_sec_perc,
                edu_post_info_tech_perc,
                ocu_sci_perc,
                ocu_mgmt_perc,
                ocu_busi_perc,
                ocu_health_perc,
                ocu_edu_perc,
                ocu_art_perc,
                ocu_trades_perc,
                ocu_natres_perc,
                ocu_man_perc)


das_census <- das %>%
  left_join(census_data_clean,
            by = c("DAUID"= "GeoUID"))

```  

* Population density. 
    * Barrington-Leigh (2017) found a U-shaped relationship between completeness 
    of roads and population density at a global scale (compared to remote 
    sensing). Densely populated places have many mappers, and sparsely populated 
    places have few roads to map.
    * Multiple studies show that rural areas may have less complete attribution, 
    since the data source is more often bulk data uploads and bots, and less 
    often local editing.
    * We expect a positive association between quality and population density.
```{r, echo = F, message = FALSE}
library(tmap)


pop_dens_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("pop_dens", style="quantile", title = "Population Density") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.25) +
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
  
  
pop_dens_map

```

* Median household income from Canada Census as a measure of economic status.
   * We expect a positive association between household income and data quality.
```{r, echo = F, message = FALSE}
library(tmap)


med_inc_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("med_hous_inc", style="quantile", title = "Median Household Income ($)") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.5) +
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
med_inc_map

```
* % of population that is a visible minority from Canada Census.
   * We expect lower data quality where there are more people who are members of 
   a visible minority group.
```{r, echo = F, message = FALSE}
library(tmap)


vis_min_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("vis_min_perc", style="quantile", title = "Visible Minorities (%)") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.5) +
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
vis_min_map

```
* Education. % of population with a post secondary degree.
* Education. % of pop with a post secondary degree in information technology.
    * We expect higher data quality where education is higher.
```{r, echo = F, message = FALSE}
library(tmap)


educ_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("edu_post_sec_perc", style="quantile", title = "Post-secondary education (%)") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.5)+
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
educ_map

```

```{r, echo = F, message = FALSE}
library(tmap)


educ_info_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("edu_post_info_tech_perc", style="quantile", title = "Post-secondary education in information technology (%)") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.5) +
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
educ_info_map

```
* Occupation % of population employed in science.
    * We expect higher data quality where more people work in science.

```{r, echo = F, message = FALSE}
library(tmap)


ocu_info_map <- tm_shape(das_census %>% filter(CSDNAME %in% "Vancouver") %>%
                           st_transform(4326)) + 
  tm_fill("ocu_sci_perc", style="quantile", title = "Science occupation (%)") +
  tm_borders(alpha = 0.1) +
  tm_layout(legend.position = c("left", "bottom"), 
            legend.title.size = 0.8,
            legend.text.size = 0.5) +
  tm_shape(reference_points_vancouver_geo) + 
  tm_dots(#col = "match_osm",
          shape = "match_osm",
          size = 0.25)
ocu_info_map

```

* The [Canadian Index of Multiple Deprivation](https://www150.statcan.gc.ca/n1/pub/45-20-0001/452000012019002-eng.htm). 
Composite indices related to the variables described above. 
  1. Residential instability (rates of home ownership, persons living alone,
  proportion of people who moved within five years, proportion of married 
  people).
  2. Economic dependency (people older than 65, people in the labour force,
  ratio of employed population, dependency ratio, people recieving gov't 
  transfer payments)
  3. Ethno-cultural composition (visible minorities, foreign born, no knowledge 
  of either official language, recent immigrants)
  4. Situational composition (indigenous population, houses needing repair, 
  population aged 25-64 without a highschool education)
```{r, echo = F, warning=F}
deprivation <- read_csv("data/deprivation/can_scores_quintiles_EN.csv") %>%
  mutate(PRCDDA = as.character(PRCDDA)) %>%
  rename(ethno_cultural = 'Ethno-cultural composition Quintiles',
         ethno_cultural_score = 'Ethno-cultural composition Scores',
         situational_vulnerability = 'Situational vulnerability Quintiles',
         situational_vulnerability_score = 'Situational vulnerability Scores',
         economic_dependency = 'Economic dependency Quintiles',
         economic_dependency_score = 'Economic dependency Scores',
         residential_instability = 'Residential instability Quintiles',
         residential_instability_score = 'Residential instability Scores') %>%
  mutate(ethno_cultural = as.factor(ethno_cultural),
         situational_vulnerability = as.factor(situational_vulnerability),
         economic_dependency = as.factor(economic_dependency),
         residential_instability = as.factor(residential_instability))

das_dep <- das %>%
  #filter(PRNAME %in% "British Columbia / Colombie-Britannique") %>%
  left_join(deprivation,
            by = c("DAUID"= "PRCDDA"))

colour_scheme <- c("#ffffb2",
                   "#fecc5c",
                   "#fd8d3c",
                   "#f03b20",
                   "#bd0026")


ethno_cultural_map <- ggmap(basemap) +
  geom_sf(data = das_dep %>% st_transform(4326), 
          aes(fill = ethno_cultural), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "YlOrBr") + 
  labs(col = "Ethno-cultural deprivation") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

situational_vulnerability_map <- ggmap(basemap) +
  geom_sf(data = das_dep %>% st_transform(4326), 
          aes(fill = situational_vulnerability), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "YlOrBr") + 
  labs(col = "Situational vulnerability") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)
economic_dependency_map <- ggmap(basemap) +
  geom_sf(data = das_dep %>% st_transform(4326), 
          aes(fill = economic_dependency), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "YlOrBr") + 
  labs(col = "Economica dependency") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

residential_instability_map <- ggmap(basemap) +
  geom_sf(data = das_dep %>% st_transform(4326), 
          aes(fill = residential_instability), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "YlOrBr") + 
  labs(col = "Residential instability") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

ethno_cultural_map
situational_vulnerability_map
economic_dependency_map
residential_instability_map
  
```
    
***Intrinsic indicators of data quality***

* count of edits within 1 km bounding box of the centroid from Ohsome API. 
   * We expect more editing will result in higher data quality. 
* count of editors within 1 km bounding box of the centroid from Ohsome API. 
   * We expect more editors will result in higher data quality. 

**Models**

* We first evaluated univariate relationships (above).
* In the first model, we evaluated the sample points alone with correct 
classification for open data and open street map as outcome variables.
* In the second model, we evaluated the number of OSM edits and editors as 
outcome variables.
* We used a Random Forests models.
* We used variable importance and partial dependence plots to understand 
relationships.

##Results

***Outcome variables***











```{r}
Can_BICS <- st_read("../osm_infrastructure/data/metrics/CAN_BICS_metric_Jan_2022.shp") %>%
  st_drop_geometry() %>%
  mutate(dauid = as.character(dauid),
         CBICS_cat = as.factor(CBICS_cat),
         ale_class = as.factor(ale_class)) %>%
  dplyr::select(dauid,
         bike_per,
         st_per,
         ale_index,
         ale_class,
         CBICS_cont,
         CBICS_cat) 

names(Can_BICS)
names(das_dep_census)

das_dep_census_cbics <- das_dep_census %>%
  left_join(Can_BICS,
            by = c("DAUID"= "dauid"))

Can_BICS_map <- ggmap(basemap) +
  geom_sf(data = das_dep_census_cbics %>% st_transform(4326), 
          aes(fill = CBICS_cat), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "PuBuGn") + 
  labs(col = "Ethno-cultural deprivation") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

Can_BICS_map

Can_ALE_map <- ggmap(basemap) +
  geom_sf(data = das_bc_census_cbics %>% st_transform(4326), 
          aes(fill = ale_class), 
          inherit.aes = FALSE,
          size = 0.25) + 
  #scale_fill_manual(values = colour_scheme) +
  scale_fill_brewer(palette = "PuBu") + 
  labs(col = "Ethno-cultural deprivation") + 
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

Can_ALE_map

```

Extract to sample points (area weighted means)

* 4 deprivation indices - done
* bike mode share
* pop dens
* Can-BICS
* Can-ALE

```{r}

ref_pts_predictors <- st_interpolate_aw(das_dep_census_cbics[c("ethno_cultural_score",
                                                       "situational_vulnerability_score",
                                                       "economic_dependency_score",
                                                       "residential_instability_score",
                                                       "pop_dens",
                                                       "bike_per",
                                                       "st_per",
                                                       "ale_index",
                                                       "CBICS_cont")],
                                reference_points %>% st_buffer(100),
                                extensive = F)

st_write(ref_pts_predictors,
         "data/ref_pts_predictors.gpkg",
         delete_dsn = T)

# combine with ref data
ref_pts_predictors_all <- cbind(reference_points,
                                ref_pts_predictors %>% st_drop_geometry()) %>%
  dplyr::select(sample_id,
                CSDNAME,
         Can_BICS_ground,
         Can_BICS_ground_class,
         Can_BICS_stratum,
         Can_BICS_stratum_class,
         Can_BICS,
         Can_BICS_class,
         match_osm,
         match_open_data,
         ethno_cultural = ethno_cultural_score,
         situational_vuln = situational_vulnerability_score,
         economic_depend = economic_dependency_score,
         residential_instability = residential_instability_score,
         pop_dens,
         bike_mode_share = bike_per,
         sustainable_mode_share = st_per,
         ale_index,
         Can_BICS_index = CBICS_cont)

```


Plot distributions
* (correct, incorrect) * (osm, open data)

```{r}
# ridge plot
ref_pts_predictors_all_long <- ref_pts_predictors_all %>%
  st_drop_geometry() %>%
  data.frame() %>%
  # filter(CSDNAME %in% "Vancouver") %>% # select geographic subset
  dplyr::select(!CSDNAME) %>%
  mutate(match_open_data = factor(match_open_data,
                               levels = c("TRUE", "FALSE")),
         match_osm = factor(match_osm,
                               levels = c("TRUE", "FALSE"))) %>%
  dplyr::select(!(Can_BICS_ground:Can_BICS_class)) %>%
  pivot_longer(
    cols = (! c(sample_id, match_open_data, match_osm)),
    names_to = "variable")

ref_pts_predictors_all_long$variable <- factor(ref_pts_predictors_all_long$variable,
                                        levels = c(
                                          "pop_dens",
                                          "bike_mode_share",
                                          "sustainable_mode_share",
                                          "ale_index",
                                          "Can_BICS_index",
                                          "ethno_cultural",
                                          "situational_vuln",
                                          "economic_depend",
                                          "residential_instability"))

library(ggridges)
colour_scheme <- c("#1f78b4", "#a6cee3")
library(ggridges)
osm_ridge <- ref_pts_predictors_all_long %>%
  ggplot(aes(
    y = fct_rev(match_osm), 
    x = value,
    fill = match_osm)) +
  scale_fill_manual(values = (colour_scheme)) +
  geom_density_ridges(alpha = 0.6,
                      scale = 3) +
  theme_ridges() +
  theme(axis.text.y=element_blank()) +
  #scale_x_continuous(limits = c(0, 5)) +
  ylab("Density") + 
  xlab("Value") +
  labs(fill = "OSM Correct") +
  facet_wrap(~(variable),
             ncol = 3,
             strip.position = "top",
             scales = "free")

osm_ridge

ggsave(osm_ridge, 
       width = 8.5,
       height = 11,
       units = "in",
       filename = "results/osm_ridge.png", dpi = 200)

colour_scheme <- c("#33a02c", "#b2df8a")

open_data_ridge <- ref_pts_predictors_all_long %>%
  ggplot(aes(
    y = fct_rev(match_open_data), 
    x = value,
    fill = match_open_data)) +
  scale_fill_manual(values = (colour_scheme)) +
  geom_density_ridges(alpha = 0.6,
                      scale = 3) +
  theme_ridges() +
  theme(axis.text.y=element_blank()) +
  #scale_x_continuous(limits = c(0, 5)) +
  ylab("Density") + 
  xlab("Value") +
  labs(fill = "Open Data Correct") +
  facet_wrap(~(variable),
             ncol = 3,
             strip.position = "top",
             scales = "free")

ggsave(open_data_ridge, 
       width = 8.5,
       height = 11,
       units = "in",
       filename = "results/open_data_ridge.png", dpi = 200)

open_data_ridge
```


Mapathon hotspots
```{r}
# select hotspots
mapathon_edits.KDE.rast[mapathon_edits.KDE.rast < 0.9] <- NA
mapathon_edits.KDE.rast[mapathon_edits.KDE.rast >= 0.9] <- 1
plot(mapathon_edits.KDE.rast)

# redefine projection to lambert, not wgs 84!
crs(mapathon_edits.KDE.rast) <- crs(reference_points)

hotspots <- rasterToPolygons(mapathon_edits.KDE.rast,
                             dissolve = T) %>%
  st_as_sf() %>%
  st_cast("POLYGON") # multipart to singlepart

hotspots$area <- st_area(hotspots)
# hotspots <- hotspots %>%
#   filter(as.numeric(area) > 1000000) # must be greater than 1 km 2

# select das touching hotspots
das_hotspots <- das_dep_census_cbics %>%
  st_intersection(hotspots)

das_dep_census_cbics$mapathon_hotspot <- das_dep_census_cbics$DAUID %in%
  das_hotspots$DAUID

das_dep_census_cbics_vanc <- das_dep_census_cbics %>%
  filter(CSDNAME %in% "Vancouver")

  mapathon_map_hotspots <- ggmap(basemap) +
    # inset_raster(as.raster(mapathon_edits.KDE.rast.geo),
    #              xmin = mapathon_edits.KDE.rast.geo@extent[1], 
    #              xmax = mapathon_edits.KDE.rast.geo@extent[2],
    #              ymin = mapathon_edits.KDE.rast.geo@extent[3], 
    #              ymax = mapathon_edits.KDE.rast.geo@extent[4]) +
    geom_sf(data = das_dep_census_cbics_vanc %>%
              filter(mapathon_hotspot), 
          aes(fill = "red"), 
          inherit.aes = FALSE,
          size = 3) + 
  labs(col = "Mapathon edits (hotspots)") +
  My_Theme +
  theme_void() +
  annotation_scale(location = "br", 
                   width_hint = 0.45,
                   pad_x = unit(-0.0001, "in")) +
  annotation_north_arrow(location = "br", which_north = "true", 
                         pad_x = unit(-0.1, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_minimal)

  mapathon_map
  
```

```{r}
  # hotspot ridge plot


das_dep_census_cbics_vanc_long <- das_dep_census_cbics_vanc %>%
  st_drop_geometry() %>%
  data.frame() %>%
    dplyr::select(
      DAUID,
      mapathon_hotspot,
      ethno_cultural = ethno_cultural_score,
      situational_vuln = situational_vulnerability_score,
      economic_depend = economic_dependency_score,
      residential_instability = residential_instability_score,
      pop_dens,
      bike_mode_share = bike_per,
      sustainable_mode_share = st_per,
      ale_index,
      Can_BICS_index = CBICS_cont) %>%
  mutate(mapathon_hotspot = factor(mapathon_hotspot,
                               levels = c("TRUE", "FALSE"))) %>%
  pivot_longer(
    cols = (! c(DAUID, mapathon_hotspot)),
    names_to = "variable")

das_dep_census_cbics_vanc_long$variable <- factor(das_dep_census_cbics_vanc_long$variable,
                                        levels = c(
                                          "pop_dens",
                                          "bike_mode_share",
                                          "sustainable_mode_share",
                                          "ale_index",
                                          "Can_BICS_index",
                                          "ethno_cultural",
                                          "situational_vuln",
                                          "economic_depend",
                                          "residential_instability"))

library(ggridges)
colour_scheme <- c("#6a3d9a", "#cab2d6")
hotspots_ridge <- das_dep_census_cbics_vanc_long %>%
  ggplot(aes(
    y = fct_rev(mapathon_hotspot), 
    x = value,
    fill = mapathon_hotspot)) +
  scale_fill_manual(values = (colour_scheme)) +
  geom_density_ridges(alpha = 0.6,
                      scale = 3) +
  theme_ridges() +
  theme(axis.text.y=element_blank()) +
  #scale_x_continuous(limits = c(0, 5)) +
  ylab("Density") + 
  xlab("Value") +
  labs(fill = "Mapathon hotspot (Vancouver)") +
  facet_wrap(~(variable),
             ncol = 3,
             strip.position = "top",
             scales = "free")

hotspots_ridge
```


